\bf Fast and Efficient Black Box Optimization using the Parameter-less Population Pyramid

### 2014-08-28 Fast and Efficient Black Box Optimization using the Parameter-less Population Pyramid/main.tex
Hill climber section and algorithm.

\subsection{Hill Climber}
~\cite{goldman:2014:p3}

@@

TODO Compare all 6 optimization techniques on time to reach global optimum.  Figure~\ref{fig-evals-to-success}.

@@

TODO Include a table of $O(N)$ seconds per evaluation for all 4 model building techniques on all 7 problems.

>>>>

\subsection{Hill Climber}
Perhaps the simplest black box search heuristic is stochastic local search, or hill climbing.
This optimization technique focuses on improving a single solution until it reaches a local
optimum. Here we use the first improvement hill climber defined by \cite{goldman:2014:p3}
and given in Figure~\ref{fig-hc}. This algorithm works by flipping each bit in a random
order, keeping modifications when fitness is improved, until doing so cannot result in
further fitness improvements.

The hill climber requires $O(1)$ operations per evaluation amortized cost. In order to
terminate, it must perform at least $N$ evaluations, where $N$ is the number of bits
in the solution, in order to ensure no single bit flip can result in a fitness improvement.
As such any operation that happens only once per search can be amortized over at least $N$
evaluations, covering the initialization of $options$.

Assuming proper hashing data structures, determining if $index \notin tried$ should require only $O(1)$
time. While this operation may be called without a directly related evaluation, amortized cost
provides an upper bound. Consider that in order to add an $index$ into $tried$ an evaluation must
be performed. Lets say this evaluation pays for two checks of that $index$, one of which was used
during this loop. If no fitness improving move is found, the next time through the loop the second check
is used. A third check is only necessary if no fitness improving move was found for any index.
This is a contradiction as the loop terminates when no improving move was found. Therefore this check
can only be called twice per evaluation.

\begin{figure}
  \begin{algorithmic}
  \Procedure{Hill-Climber}{}
    \State $options \leftarrow [0 \dots N-1]$
    \State $tried \leftarrow \emptyset$
    \While{$|tried| < |options|$}
      \State $shuffle(options)$
      \ForAll{$index \in options$}
        \If{$index \notin tried$}
          \State Flip bit $index$ in solution
          \If{solution's fitness increased}
            \State $tried \leftarrow \emptyset$
          \Else
            \State Revert change
          \EndIf
          \State $tried \leftarrow tried \cup \{index\}$
        \EndIf
      \EndFor
    \EndWhile
  \EndProcedure
\end{algorithmic}
  \caption{Hill climbing algorithm used to improve randomly generated solutions until no single
           bit change results in a fitness improvement.}
  \label{fig-hc}
\end{figure}

@@

TODO Compare all 6 optimization techniques on evaluations to reach global optimum.  Figure~\ref{fig-evals-to-success}.

@@

TODO Include a table of $O(N)$ seconds per evaluation and total seconds for all 4 model building techniques on all 7 problems.
